---
title: "Visualizing Financial Inclusion"
author: "Marissa Block"
output:
  html_document:    
    code_folding: hide
    css: style.css
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


```{r}
library(tidyverse)
library(here)
library(treemapify)
library(RColorBrewer)
library(haven)
library(readxl)
library(sf)
library(extrafont)
library(ggrepel)
library(maptools)

# Load data
findex <- read_csv("../fininc-data/clean/clean.findex.csv")
imf <- read_csv("../fininc-data/clean/clean.imf.csv")
finac.k.15 <- read_csv("../fininc-data/clean/clean.finac.k.15.csv")
finac.k.13 <- read_csv("../fininc-data/clean/clean.finac.k.13.csv")
finac.k.06 <- read_csv("../fininc-data/clean/clean.finac.k.06.csv")
wdi <- read_csv("../fininc-data/clean/clean.wdi.csv")
pop <- read_csv("../fininc-data/clean/clean.pop.csv")


# Get geospatial data
kenya <- read_sf(dsn="../fininc-data/ken_admbnda_adm1_iebc_20180607", layer = "ken_admbnda_adm1_iebc_20180607")
  
  # Cut excess variables
  kenya <- kenya %>%
    select(Shape_Leng, Shape_Area, ADM1_EN, geometry) %>%
    rename(county = "ADM1_EN")

  # Project data
  kenya <- st_transform(kenya, 21097)
  

```


# Financial Access Survey, IMF
Technology has been instrumental in improving access to finance, namely mobile money and digital payments. East Africa, and Kenya in particular, have been at the forefront of mobile money. Mobile money allows for easy and safe transfers of resources among households and individuals and has significantly reduced transaction costs.

```{r, fig.height = 5, fig.width = 7, fig.align = 'center'}
imf %>%
  filter(!is.na(i_mob_transactions_value_GDP), year>2011) %>%
  group_by(Region, year) %>%
  summarize(mean = mean(i_mob_transactions_value_GDP, na.rm = TRUE)) %>%
  mutate(Region2 = factor(Region, levels = c("Sub-Saharan Africa", "East Asia & Pacific", "South Asia", "Europe & Central Asia", "Latin America & Caribbean",
                                            "Middle East & North Africa"))) %>%
  ggplot(aes(x = year, y = mean, color = Region2)) +
  geom_line(size = 0.5) +
  labs(
    title = "Mobile money useage has grown significantly in Sub-Saharan Africa",
    subtitle = "Measured as average value of mobile money transactions (% of GDP)",
    caption = "Source: IMF Financial Access Survey",
    x = "Year",
    y = "Percent of GDP"
  ) +
  scale_x_continuous(breaks = c(2012, 2013, 2014, 2015, 2016, 2017)) +
  scale_color_brewer(name = "Region", palette = "Dark2") +
  geom_point() +
  theme(text = element_text(family = "Avenir Next"),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(face = "italic"),
        #panel.background = element_rect(fill = "white", color = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = c(0.05, 0.97),
        legend.justification = c("left", "top"),
        legend.box.just = "left",
        legend.margin = margin(6, 6, 6, 6),
        legend.box.background = element_rect())

```

Finance may be considered a constraint to economic growth if it's supply is constrained relative to demand. One means of understanding this is by analyzing the cost of finance, as measured by the real interest rate. A high interest rate, indicates a high cost of finance. A high cost of finance could be the result of low savings or costly intermediation. If low savings is the culprit, then we would expect that countries with a high interest rate also have a low savings, which is measured here as the number of depositors with commerical banks per 1,000 adults. A large number of depositors may indicate a large portion of savings.

Below, we see Madagascar has a very high interest rate and a small number of depositors. Thus, one might conclude that low savings is causing the high cost of finance. Brazil also has a very high interest rate, but the number of depositors is not quite as small as Madagascar. In contrast, the interest rate in Singaporeis not very high, and there are a large number of depositors, indicating that finance is not a constraint to the economy.

```{r, fig.height = 5, fig.width = 9, fig.align = 'center'}
  
  # Create dataframe of labels - MDG, TLS, BRA, QAT, SGP, KWT, BRN, ARG, SUR
  labels.bubble <-   wdi %>%
    filter(year==2016, cntry.code!="SSD") %>%
    select(cntry.code, year, ny.gdp.pcap.kd, fr.inr.rinr) %>%
    left_join(imf) %>%
    select(cntry.code, economy, year, ny.gdp.pcap.kd, fr.inr.rinr, i_depositors_A1_pop, i_deposit_acc_A1_pop, Region) %>%
    filter(!is.na(fr.inr.rinr)) %>%
    filter(cntry.code %in% c("MDG", "TLS", "BRA", "QAT", "SGP", "KWT", "BRN", "ARG", "SUR", "BWA"))
    
  
  wdi %>%
    filter(year==2016) %>%
    select(cntry.code, year, ny.gdp.pcap.kd, fr.inr.rinr) %>%
    left_join(imf) %>%
    select(cntry.code, year, ny.gdp.pcap.kd, fr.inr.rinr, i_depositors_A1_pop, i_deposit_acc_A1_pop, Region) %>%
    filter(!is.na(fr.inr.rinr) & !is.na(ny.gdp.pcap.kd) & !is.na(i_depositors_A1_pop)) %>%
  ggplot(aes(x = ny.gdp.pcap.kd, y = fr.inr.rinr, size = i_depositors_A1_pop, color = Region)) +
    geom_point(alpha = 0.7) +
   geom_label_repel(data = labels.bubble, aes(label = economy), family = "Avenir Next", size = 3,
                   show.legend = FALSE, angle = 30, nudge_x = 0.2, box.padding = 0.5) +
    #geom_label(aes(label = cntry.code)) +
    scale_color_brewer(palette = "Dark2") +
    labs(
      title = "The Cost of Formal Finance",
      subtitle = "As measured by the real interest rate, GDP per capita",
      caption = "Source: World Development Indicators, IMF Financial Access Survey, 2016",
      y = "Real Interest Rate",
      x = "GDP per capita, (constant 2010 US$)"
    ) +
    theme(text = element_text(family = "Avenir Next"),
          plot.title = element_text(face = "bold"),
          plot.subtitle = element_text(face = "italic")) +
    annotate("rect", xmin = 0, xmax = 20000, ymin = 38, ymax = 55, alpha = 0.15) +
    annotate("text", x = 17000, y = 50, label = "High cost of finance", family = "Avenir Next") +
    annotate("rect", xmin = 1000, xmax = 16000, ymin = -18, ymax = -2, alpha = 0.15) +
    annotate("text", x = 18000, y = -15, label = "Low cost of finance", family = "Avenir Next") +
    scale_size_continuous(name = "# of Depositors per 1,000 adults")
  
```


# The Global Findex
```{r, fig.width = 9, fig.height = 7, fig.align = 'center'}
# Get map of Africa
data(wrld_simpl)
afr <- wrld_simpl[wrld_simpl$REGION==2,]
afr <- st_as_sf(afr)
afr <- rename(afr, cntry.code = "ISO3")

# Merge with Findex
data <- afr %>%
  left_join(findex, by = "cntry.code") %>%
  filter(year==2017) %>%
  select(cntry.code, geometry, account.t.d, country) %>%
  st_as_sf()

# Make labels
labels <- data %>%
  cbind(st_coordinates(st_centroid(data))) %>%
  filter(country=="Kenya" | country=="Namibia" | country=="South Africa" | country=="Central African Republic" | country=="Niger")
labels$country[labels$country=="Central African Republic"] <- "CAR"

# Map
  ggplot() +
  geom_sf(data = afr) +
    geom_sf(data = data, aes(fill = account.t.d)) +
    geom_label_repel(data = labels, aes(X, Y, label = country), family = "Avenir Next", size = 3) +
    coord_sf(datum=NA) +
    scale_fill_distiller(palette = "Greens", direction = 1, name = "Proportion of \nrespondents") +
    theme(panel.background = element_blank(),
          text = element_text(family = "Avenir Next"),
          plot.title = element_text(face = "bold"),
          plot.subtitle = element_text(face = "italic"),
          axis.title = element_blank(),
          legend.position = "bottom", legend.direction = "horizontal",
          ) +
    annotate("text", x = -5, y = -12, label = "Gray represents \nno data", family = "Avenir Next", size = 4) +
    labs(
      title = "Kenya, Namibia, and South Africa have the highest \nproportion of the population having a financial account",
      subtitle = "Whereas financial access is low in Niger and \nthe Central African Republic (CAR)",
      caption = "Source: Global Findex, 2017")
```


## Mobile Money in Africa

```{r, fig.width = 10, fig.height = 7, fig.align = 'center'}
findex %>%
  filter(year == 2017, region=="Sub-Saharan Africa (excluding high income)") %>%
  select(country, mobileaccount.t.d.1, mobileaccount.t.d.2) %>%
  filter(!is.na(mobileaccount.t.d.1)) %>%
  arrange(mobileaccount.t.d.2) %>%
  mutate(country = factor(country, country), mobileaccount.t.d.1 = mobileaccount.t.d.1*100, 
         mobileaccount.t.d.2 = mobileaccount.t.d.2*100) %>%
  ggplot() +
  geom_segment(aes(x=country, xend = country, y = mobileaccount.t.d.2,
                   yend = mobileaccount.t.d.1), color = "black") +
  geom_point(aes(x = country, y = mobileaccount.t.d.2), color = "#7570b3", size = 2) +
  geom_point(aes(x = country, y = mobileaccount.t.d.1), color = "#1b9e77", size = 2) +
  geom_text(aes(x = "Kenya", y = 65, label = "Female"), nudge_x = -0.7, nudge_y = 0.9,
            color = "black", size = 3, family = "Avenir Next", angle = 20) +
  geom_text(aes(x = "Kenya", y = 78, label = "Male"), nudge_x = -0.7,
            color = "black", size = 3, family = "Avenir Next", angle = 20) +
  coord_flip() +
  labs(
    title = "While Africa leads the way in mobile money access, in almost all countries, a higher proportion 
of males have mobile money accounts",
    subtitle = "Gender Difference in Mobile Money Access, 2017",
    caption = "Note: Excludes South Sudan and Central Africa Republic, for which there is no data.
    Source: Global Findex, 2017",
    y = "% of respondents with mobile money accounts (Age +15)",
    x = NULL
  ) +
  scale_y_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80)) +
  theme(text = element_text(family = "Avenir Next"),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(face = "italic"),
        panel.grid.minor.y =element_blank(),
        panel.grid.major.y = element_blank())

```
Kenya leads the way in mobile money, while in some countries, like Ethiopia, mobile money has not taken off. 


## A Closer Look at Kenya

The following charts utilize the 2015-16 Financial Access Survey from Kenya.

In general, Kenya does well on providing access to finance. However, individuals who have never banked represent the largest proportion in the 2015-16 Finance Access Survey.
```{r, fig.width = 8, fig.height = 7, fig.align = 'center'}
  ## Dot Density of banked and unbanked

# Create county labels
  labels3 <- kenya %>%
  cbind(st_coordinates(st_centroid(kenya))) %>%
  filter(county=="Mandera" | county=="Garissa" | county=="Wajir")

  # No bank
  nobank <- finac.k.15 %>%
    filter(bank_usage==3) %>%
    group_by(county) %>%
    summarize(n = n())
  
  # Used to bank
  prevbank <- finac.k.15 %>%
    filter(bank_usage==2) %>%
    group_by(county) %>%
    summarize(n2 = n())
  
  # Merge no bank + used to bank
  bank.usage <- nobank %>%
    left_join(prevbank, by = "county")
  
  
  sf <- finac.k.15 %>%
    filter(bank_usage==1) %>%
    group_by(county) %>%
    summarize(n3 = n()) %>%
    left_join(bank.usage, by = "county") %>%
    left_join(kenya, by = "county") %>%
    st_as_sf() %>%
    select(county, n3, n, n2, geometry)

  random_round <- function(x) {
    v=as.integer(x)
    r=x-v
    test=runif(length(r), 0.0, 1.0)
    add=rep(as.integer(0),length(r))
    add[r>test] <- as.integer(1)
    value=v+add
    ifelse(is.na(value) | value<0,0,value)
    return(value)
  }
  
  num_dots <- as.data.frame(sf) %>%
    select(n3, n, n2) %>%
    mutate_all(funs(./1)) %>%
    mutate_all(random_round)
  
  sf_dots <- map_df(names(num_dots),
                    ~st_sample(sf, size = num_dots[,.x], type = "random") %>%
                      st_cast("POINT") %>%
                      st_coordinates() %>%
                      as_tibble() %>%
                      setNames(c("lon", "lat")) %>%
                      mutate(mob = .x)
  ) %>%
    slice(sample(1:n()))
  
  ggplot() +
    geom_sf(data = sf, fill = "transparent", color = "white") +
    geom_point(data = sf_dots, aes(lon, lat, color = mob), size = 1, alpha = 0.7) +
    geom_label(data = labels3, aes(X, Y, label = county), size = 3, family = "Avenir Next") +
    coord_sf(crs = 21097, datum = NA) +
    scale_color_brewer(name = NULL, labels = c("Never banked", "Previously banked", "Currently banked"), palette = "Dark2") +
    labs(
      x = NULL,
      y = NULL,
      title = "Individuals who have never banked dominate",
      subtitle = "Some counties in the northeast tend to have more unbanked individuals",
      caption = "Source: Financial Access Survey 2015-16, Kenya"
    ) +
    theme(text = element_text(family = "Avenir Next", color = "white"),
          plot.title = element_text(face = "bold"),
          plot.subtitle = element_text(face = "italic"),
          legend.position = "bottom", legend.direction = "horizontal",
          aspect.ratio = 1.05,
          plot.background = element_rect(fill = "#212121", color = NA), 
          panel.background = element_rect(fill = "#212121", color = NA),
          legend.background = element_rect(fill = "#212121", color = NA),
          legend.key = element_rect(fill = "#212121", color = NA),
          legend.text = element_text(size = 10),
    )
  

```



Access to finance for entrepreneurs is a key ingredient to economic growth. When small-to-medium enterprises, which are characteristic of Kenya, can access financial services, these business can save, invest, and ultimately grow. Yet, surprisingly, few business owners use these services.

```{r,fig.height = 6, fig.width = 5, fig.align = 'center'}
finac.k.15 %>%
  filter(!is.na(l11_1)) %>%
  group_by(l11_1) %>%
  mutate(bank = "Account at a bank") %>%
  ggplot(aes(x = as.factor(bank), fill = as.factor(l11_1))) +
  geom_bar(position = "dodge", alpha = 0.8) +
  scale_fill_brewer(palette = "Dark2", name = NULL, labels = c("Currently use", "Used to use", "Never used")) +
  labs(
    title = "An overwhelming majority of Kenyan business-owners \nhave never used a bank account",
    subtitle = "At least for business purposes",
    caption = "Source: Financial Access Survey, 2015-2016, Kenya",
    x = NULL,
    y = NULL
  ) +
  theme(text = element_text(family = "Avenir Next"),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(face = "italic"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.position = c(0.05, 0.97),
        legend.justification = c("left", "top"),
        legend.box.just = "left",
        legend.margin = margin(6, 6, 6, 6),
        legend.box.background = element_rect())
```

Rather than taking out loans, Kenyan business owners tend to utilize their own personal savings to start their business. This could be a result of a number of issues, including high interest rates or the inability of the financial sector to appropriately access risk.

```{r, fig.align = 'center'}
# First, create dataframe of labels
label <- finac.k.15 %>%
  filter(l5_1!="NA", l5_1<995) %>%
  mutate(type = factor(l5_1, labels = c("Commercial bank loan", "Microfinance loan",
                                        "SACCO loan", "Money lender", "Chama (Savings group)", 
                                        "Loan from family/friends", "Gift from family/friends", "Income from another business", "Sale of assets", 
                                        "Own savings", "Inherited", "Government fund", "Other", "Don't know", "Mobile" ))) %>%
  group_by(type, gender_of_respondent) %>%
  summarize(n = n()) %>%
  summarize(total1 = sum(n[gender_of_respondent==1]), total2 = sum(n[gender_of_respondent==2])) %>%
  mutate(male = sum(total1), female = sum(total2), pct.m = (total1/male)*100, pct.f = (total2/female)*100) %>%
  gather(pct.f, pct.m, key = "Sex", value = "pct") %>%
  mutate(label = paste0(type, ": ", round(pct, 1), "%")) %>%
  select(type, Sex, label)
  label$Sex[label$Sex=="pct.f"] <- 2
  label$Sex[label$Sex=="pct.m"] <- 1
  label$Sex<- as.numeric(label$Sex)
# Now, get dataframe of counts and merge in labels
finac.k.15 %>%
  filter(l5_1!="NA", l5_1<995) %>%
  rename(Sex = "gender_of_respondent") %>%
  mutate(type = factor(l5_1, labels = c("Commercial bank loan", "Microfinance loan",
                                        "SACCO loan", "Money lender", "Chama (Savings group)", 
                                        "Loan from family/friends", "Gift from family/friends", "Income from another business", "Sale of assets", 
                                        "Own savings", "Inherited", "Government fund", "Other", "Don't know", "Mobile" ))) %>%
  group_by(type, Sex) %>%
  summarize(n = n()) %>%
  left_join(label) %>%
  mutate(Sex2 = factor(Sex, levels = c("1", "2"),
                       labels = c("Male", "Female"))) %>%
  ggplot(aes(area = n, fill = Sex2, label = label, subgroup = Sex2)) +
  geom_treemap(fill = "white") +
  geom_treemap(aes(alpha = n)) +
  geom_treemap_subgroup_border(color = "white") +
  geom_treemap_subgroup_text(color = "black", place = "bottomleft", family = "Avenir Next", fontface = "italic") +
  geom_treemap_text(aes(label = label),
                    place = "center",
                    grow = F,
                    color = "black",
                    min.size = 1,
                    reflow = TRUE,
                    family = "Avenir Next"
  ) +
  scale_fill_brewer(palette = "Dark2") +
  theme(legend.position = "none") +
  labs(title = "Many business owners user their own savings as start-up capital \n rather than loans",
       subtitle = "Females also receive more gifts from family or friends.",
       caption = "Source: Financial Access Survey, 2015-2016, Kenya"
  ) +
  theme(text = element_text(family = "Avenir Next"),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(face = "italic"))
```

Looking beyond business owners, Kenyans utilize all sorts of financial services. Although many business-owners do not use bank accounts for business purposes, most individuals living in urban areas do use bank accounts. However, the difference in access between urban males and females is significantly different, as is the difference between rural males and females.

```{r, fig.align = 'center'}

finac.k.15 %>%
  group_by(cluster_type, gender_of_respondent) %>%
  summarize(Bank = mean(e4_1, na.rm = TRUE)*100,
            SACCO = mean(e4_3, na.rm = TRUE)*100,
            Microfinance = mean(e4_4, na.rm = TRUE)*100,
            Savings = mean(e4_5, na.rm = TRUE)*100,
            Loans = mean(e4_7, na.rm = TRUE)*100,
            Insurance = mean(e4_9, na.rm = TRUE)*100) %>%
  gather("Bank", "SACCO", "Microfinance", "Savings", "Loans", "Insurance", key = "product", value = "proportion") %>%
  unite(area_sex, cluster_type, gender_of_respondent) %>%
  mutate(area_sex = factor(area_sex, levels = c("1_1", "1_2", "2_1", "2_2"),
                           labels = c("Rural male", "Rural female", "Urban male", "Urban female"))) %>%
  mutate(product = factor(product, levels = c("Savings", "Bank", "Loans", "Insurance", "SACCO", "Microfinance"))) %>%
  ggplot(aes(x = product, y = area_sex, fill = proportion)) +
  geom_tile(color = "black", size = 0.25) +
  geom_text(aes(label = paste0(round(proportion,1), "%"), family = "Avenir Next", fontface = "bold")) +
  scale_fill_gradient2(name = "Proportion", low = "#e8f5f1", mid = "#76c4ad", high = "#1b9e77", guide = "colorbar", midpoint = 40) +
  #scale_fill_brewer(palette = "Dark2") +
  labs(
    title = "Most Kenyans utilize savings accounts, \n regardless of sex or location",
    subtitle = "The proportion of individuals with particular financial products by sex and area",
    caption = "Source: Financial Access Survey 2015-16, Kenya",
    y = NULL,
    x = NULL
  ) +
  theme(text = element_text(family = "Avenir Next"),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(face = "italic"),
        panel.background = element_rect(fill = "white"))
```

Finally, how has financial product usage changed over time? Both savings and bank usage have increased significantly. Microfinance, once a touted solution, has seen limited change.

```{r, fig.height = 7.5, fig.width = 5, fig.align = 'center'}
## Slope graph
finac.k.06 <- mutate(finac.k.06, year = 2006)
finac.k.06 <- rename(finac.k.06, bank_usage = "A4", savings_usage = "X_savings", credit_usage = "X_credit", mfi_usage = "X_cur_mfi", insurance_usage = "X_useins")
finac.k.15 <- mutate(finac.k.15, year = 2015)


y2006 <- finac.k.06 %>%
  select(year, bank_usage, savings_usage, credit_usage, mfi_usage, insurance_usage)
y2015 <- finac.k.15 %>%
  select(year, bank_usage, savings_usage, credit_usage, mfi_usage, insurance_usage)

combine <-bind_rows(y2006, y2015)
combine <- combine %>%
  rename(Banks = "bank_usage", Savings = "savings_usage", Credit = "credit_usage", Microfinance = "mfi_usage", Insurance = "insurance_usage") %>%
  gather(Banks, Savings, Credit, Microfinance, Insurance, key = "type", value = "value") %>%
  group_by(year, type) %>%
  summarize(total1 = sum(value[value==1]), total2 = n(), pct = (total1/total2)*100)

  ggplot(data = combine, aes(x = year, y = pct, group = type)) +
    geom_line(aes(color = type, alpha = 1), size = 2) +
    geom_point(aes(color = type, alpha = 1), size = 4) +
    scale_x_continuous(name = "Year", position = "top", breaks = c(2006, 2015)) +
    scale_color_brewer(palette = "Dark2") +
    geom_text_repel(data = combine %>% filter(year==2006), aes(label = paste0(type, ":", round(pct, 2), "%")),
              hjust = -0.1,
              vjust = -1,
              family = "Avenir Next",
              size = 4) +
    geom_text_repel(data = combine %>% filter(year==2015), aes(label = paste0(type, ":", round(pct, 2), "%")),
              hjust = 1,
              vjust = -1,
              family = "Avenir Next",
              size = 4) +
    labs(
      title = "Commercial banks and savings products have seen the \n largest increase in usage from 2006 to 2015",
      subtitle = "Credit and microfinance have hardly changed, \n while insurance has decreased slightly",
      caption = "Note: Individuals may have more than one financial product, \n so these numbers do not sum to 100. \n Source: Financial Access Survey, Kenya, 2006 and 2015"
    ) +
    theme(text = element_text(family = "Avenir Next"),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(face = "italic"),
        legend.position = "none",
        panel.border = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.ticks = element_blank())
```

